Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2022-04-05 paper/manuscript.Rmd
Update figures

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage[table]{xcolor}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE
)
```

@@

fig1_2 <- get_plot_rich_bm(
  predict_plot = pred_bm_rich_mono_stable_trends,
  get_list = TRUE,
  rm_legend = FALSE,
  bm_x = "log_bm_std",
  bm_y = c("ct_ff", "w_trph_lvl_avg"), #, "piel_nind"
  rich_y = c("ct_ff", "w_trph_lvl_avg", "log_bm_std", #, "piel_nind"
  rich_x = "log_rich_std"
  )
p_bm <- fig1_2$bm
names(p_bm) <- c("ct_ff", "w_trph_lvl_avg")
#plot_grid(plotlist = p_bm)

p_sp_bm <- sp_fig1_2_bm_rich_mono_stable_trends$bm
names(p_sp_bm) <- c("log_rich_std", "ct_ff", "w_trph_lvl_avg")
```

@@

ggMarginal(p_bef, type = "histogram")
```

@@

```{r}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~.x +
    theme(
      axis.text = element_text(size = 6),
      axis.title = element_text(size = 8)
      )
  )
fig2$bm <- map2(fig2$bm, names(fig2$rich),
  ~ .x + theme(
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)
    ) +
  labs(
    y = paste0(var_replacement()[.y],
      "\ntrends (per year)"),
    x = lab_log_bm
  ) +
  scale_xy_log_trends(breaks = c(-.40, -.20, 0, .20, .40), axis = "x")
)
# Delete redondant legend
fig2$bm[1:2] <- map(fig2$bm[1:2],
  ~ .x + theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )
  )

fig2$rich <- map(fig2$rich,
  ~ .x + theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  ) +
  labs(x = lab_log_rich) +
  scale_xy_log_trends(breaks = c(-.10, -.050, 0, .050, .10), axis = "x")
  )
fig2$rich[1:2] <- map(fig2$rich[1:2],
  ~ .x + theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )
  )

@@

```{r}
# Add histograms
fig2$bm[[1]] <- ggMarginal(fig2$bm[[1]], type = "histogram", margins = "x")
fig2$rich[[1]] <- ggMarginal(fig2$rich[[1]], type = "histogram", margins = "both")
fig2$rich[2:3] <- map(fig2$rich[2:3], ~ggMarginal(.x, type = "histogram", margins = "y"))
```

```{r bmr, out.width="80%", fig.height=7, fig.cap=fig_bm_rich_cap}
plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("A", "B", "C"),
    hjust = -5.5, vjust = 1.75
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("D", "E", "F"),
    hjust = -2, vjust = 1.75
    ),
  ncol = 2,
  rel_widths = c(1, 1)
)
```

```{r}
std_cap <- paste0(
  "Assessment of the space-for-time hypothesis. Effect of community biomass (top
  panel) and species richness (bottom panel) on food-web structure for the
temporal and spatial models. The bar tops display the standardized slope coefficients
from the models ($r_\\delta$), the lines around bar tops display the standard errors.")
```

```{r std, fig.cap=std_cap, fig.width=6}
#loadd(fig_std_coef)
std_table_bm_rich_mono_stable_trends %>%
  filter(
    term %in% c("log_rich_std", "log_bm_std"),
    !response %in% "piel_nind") %>%
  mutate_at(vars(term, response), ~str_replace_all(., var_replacement(
        several_lines = TRUE))) %>%
  mutate(
    type = str_to_sentence(type),
    type = factor(type, levels = c("Temporal", "Spatial"))) %>%
  ggplot(aes(y = std_estimate, x = response, fill = type)) +
  viridis::scale_fill_viridis(begin = 0.5, end = 1, discrete = TRUE) +
  facet_grid(rows = vars(term)) +
  geom_col(position = position_dodge(width = .9)) +
  geom_linerange(aes(ymin = std_estimate - stdse, ymax = std_estimate + stdse),
    position = position_dodge(width = .9)) +
  labs(
    x = "Response variables",
    y = expression(paste("Standardized slope coefficients ", r[delta])),
    fill = "Model") +
  theme(legend.position = c(0.6, .05),
    legend.direction = "horizontal",
    text = element_text(size = 8), axis.text = element_text(size = 8)) +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = "Predictive variable",
      breaks = NULL, labels = NULL))
```

>>>>

---
title: "Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs"
author: Alain Danet, Loubna El Madouri, Maud Mouchet, Elisa Thébault and Colin Fontaine
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
  bookdown::word_document2:
fontsize: 12pt
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{multirow}
   - \usepackage{booktabs}
   - \usepackage{makecell}
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
number_sections: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

@@

fig1_2 <- get_plot_rich_bm(
  predict_plot = pred_bm_rich_mono_stable_trends,
  get_list = TRUE,
  rm_legend = FALSE,
  bm_x = "log_bm_std",
  bm_y = c("ct_ff", "w_trph_lvl_avg"), #, "piel_nind"
  rich_y = c("ct_ff", "w_trph_lvl_avg", "log_bm_std"), #, "piel_nind"
  rich_x = "log_rich_std"
  )
p_bm <- fig1_2$bm

p_sp_bm <- sp_fig1_2_bm_rich_mono_stable_trends$bm[c("log_rich_std", "ct_ff", "w_trph_lvl_avg")]
```

@@

@@

```{r}
fig2 <- fig1_2
fig2$rich$log_bm_std <- NULL
fig2$rich <- map(fig2$rich,
  ~.x +
    theme(
      axis.text = element_text(size = 6),
      axis.title = element_text(size = 8)
      )
  )
fig2$bm <- map2(fig2$bm, names(fig2$rich),
  ~ .x + theme(
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 8)
    ) +
  labs(
    y = paste0(var_replacement()[.y],
      "\ntrends (per year)"),
    x = lab_log_bm
  ) +
  scale_xy_log_trends(breaks = c(-.40, -.20, 0, .20, .40), axis = "x")
)
# Delete redondant legend
fig2$bm[[1]] <- fig2$bm[[1]] + theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  )

fig2$rich <- map(fig2$rich,
  ~ .x + theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  ) +
  labs(x = lab_log_rich) +
  scale_xy_log_trends(breaks = c(-.10, -.050, 0, .050, .10), axis = "x")
  )
fig2$rich[[1]] <- fig2$rich[[1]] + theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  )

@@

```{r}
# Add histograms
p_bef2 <- p_bef +
  theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
  ) 
p_bef_hist <- ggMarginal(p_bef2, type = "histogram")
#fig2$bm[[1]] <- ggMarginal(fig2$bm[[1]], type = "histogram", margins = "x")
fig2$rich[[1]] <- ggMarginal(fig2$rich[[1]], type = "histogram", margins = "y")
fig2$rich[[2]] <- ggMarginal(fig2$rich[[2]], type = "histogram", margins = "y")
```

```{r}
sem_cap <- "Structural equation models on the temporal trends of the food-web
structure. The path display the standardized slopes and their associated 95%
  confidence intervals."
```

```{r sem, fig.cap = sem_cap}
ti <- magick::image_read(here::here("figures", "sem_tps_avg_weight_ok.svg"),  density = 300)
ggdraw() +
  draw_image(ti,
    x = 0, y = 0, hjust = 0, vjust = 0, width = .9, height = 1)
```

```{r bmr, out.width="100%", fig.height=7, fig.cap=fig_bm_rich_cap}
p <- draw_image(here::here("figures", "sem_tps_avg_weight_ok.pdf"),
  x = 1, y = 1, hjust = 1, vjust = 1, width = 0.13, height = 0.2)
biplot_model <- plot_grid(
  plot_grid(, p_bef_hist,
    labels = c("", "A"), ncol = 2,
    rel_widths = c(.62, 1)
    ),
  plot_grid(
  plot_grid(plotlist = fig2$bm,
    ncol = 1, labels = c("B", "D"),
    hjust = -5.5, vjust = 1.75
    ),
  plot_grid(plotlist = fig2$rich,
    ncol = 1, labels = c("C", "E"),
    hjust = -2, vjust = 1.75
    ),
  ncol = 2,
  rel_widths = c(.85, 1)
  ),
  nrow = 2,
  rel_heights = c(.6, 1)
)
save_plot(
  filename = here::here("figures", "biplot_model.png"),
  biplot_model)

biplot_model
```

```{r}
std_cap <- paste0(
  "Assessment of the space-for-time hypothesis. Standardized coefficients
  ($r_\\delta$) derived from the spatial and temporal Structural Equation
  Models. The left, middle and right panels display respectively the direct,
  indirect and total effects of community biomass and species richness on the
  Average trophic level (top panel), the connectance (middle panel) and the
  community biomass (bottom panel) The lines display the confidence intervals at
  95%.")

```{r std-old, eval = FALSE, fig.cap=std_cap, fig.width=6}
#loadd(fig_std_coef)
std_table_bm_rich_mono_stable_trends %>%
  filter(
    term %in% c("log_rich_std", "log_bm_std"),
    !response %in% "piel_nind") %>%
  mutate_at(vars(term, response), ~str_replace_all(., var_replacement(
        several_lines = TRUE))) %>%
  mutate(
    type = str_to_sentence(type),
    type = factor(type, levels = c("Temporal", "Spatial"))) %>%
  ggplot(aes(y = std_estimate, x = response, fill = type)) +
  viridis::scale_fill_viridis(begin = 0.5, end = 1, discrete = TRUE) +
  facet_grid(rows = vars(term)) +
  geom_col(position = position_dodge(width = .9)) +
  geom_linerange(aes(ymin = std_estimate - stdse, ymax = std_estimate + stdse),
    position = position_dodge(width = .9)) +
  labs(
    x = "Response variables",
    y = expression(paste("Standardized slope coefficients ", r[delta])),
    fill = "Model") +
  theme(legend.position = c(0.6, .05),
    legend.direction = "horizontal",
    text = element_text(size = 8), axis.text = element_text(size = 8)) +
  scale_y_continuous(sec.axis = sec_axis(~ ., name = "Predictive variable",
      breaks = NULL, labels = NULL))
```

```{r std, fig.cap=std_cap, fig.width=6, fig.height = 9}
load(here::here("data", "semeff.rda"))
load(here::here("data", "sp_semeff.rda"))
sp_semeff_tab <- from_semEff_to_table(sp_semeff)
semeff_tab <- from_semEff_to_table(semeff)

semeff_tot <- rbind(
  tibble(semeff_tab, model = "temporal"),
  tibble(sp_semeff_tab, model = "spatial")
  ) %>%
filter(
  !predictor %in% c("log_RC1", "log_RC2"),
  effect_type != "mediators"
)

semeff_tot %>%
  mutate(
    response = var_replacement()[response],
    predictor = var_replacement()[predictor]
    ) %>%
  ggplot(aes(x = predictor, y = effect, color = model)) +
  geom_hline(yintercept = 0) +
  geom_pointrange(aes(ymin = lower_ci, ymax = upper_ci), position = position_dodge(width = 0.4)) +
  facet_grid(rows = vars(response), cols = vars(effect_type)) +
  labs(
    x = "Predictor",
    y = expression(paste("Standardized slope coefficients ", r[delta])),
    color = "Model") +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```
